Supplemental Information C: Power Simulations

### 2023-06-01 Supplemental-Material-B-Power-Simulation_between.Rmd
supplemental doc init

# Install and Load Packages
# !IMPORTANT!
# BEFORE FIRST RENDER:
# To install all relevant packages please run "renv::restore()" (or renv::init() and then initiate from lockfile) in the console before the first use to ensure that all packages are using the correct version.
# to store the packages in a contained library within the project folder: renv::settings$use.cache(FALSE) and add 'RENV_CONFIG_SANDBOX_ENABLED = FALSE' to an '.Renviron' file
lib <- c(
  "rmarkdown",
  "knitr",
  "remedy",
  "bookdown",
  "brms",
  "psych",
  "ggplot2",
  "ggthemes",
  # "haven",
  "RColorBrewer",
  # "plotly",
  "gridExtra",
  "ggpattern",
  "ggridges",
  "binom",
  "iterators",
  "pbkrtest",
  "plotrix",
  "RLRsim",
  "stats",
  "methods",
  "utils",
  "graphics",
  "grDevices",
  "car",
  "testthat",
  "simr",
  "data.table",
  "lme4",
  "lmerTest",
  # "nlme",
  "jtools",
  # "gtsummary",
  "sessioninfo",
  # "tibble",
  "pander",
  # "devtools",
  # "mada",
  #"tidyr",
  #"tidyverse",
  "plyr",
  "dplyr",
  # "Hmisc",
  "kableExtra",
  # "papaja",
  "stringr"#,
  # "stringi",
  # "reshape2",
  # "lubridate",
  # "purrr",
  # "metafor"
)
invisible(lapply(lib, library, character.only = TRUE))
rm(lib)

@@

@@

load("data/wrangled.RData")
```

@@

# TBD
```

@@

# run regresion 
lmer_quality_contemporary <-
  lmer(
    quality_overall_1 ~ 
      keymotive_fulfillemt_1_cwc + (1 + keymotive_fulfillemt_1_cwc | PID),
    data = dtWorkerSupp$workerOutWithinBetween
  )
lmer_attitude_contemporary <-
  lmer(
    thermometerDutch_1 ~ 
      keymotive_fulfillemt_1_cwc + (1 + keymotive_fulfillemt_1_cwc | PID),
    data = dtWorkerSupp$workerOutWithinBetween
  )
lmer_attitude_need_quality_contemporary <-
  lmer(
    thermometerDutch_1 ~
      keymotive_fulfillemt_1_cwc + quality_overall_1_cwc + (1 + keymotive_fulfillemt_1_cwc + quality_overall_1_cwc | PID),
    data = dtWorkerSupp$workerOutWithinBetween
  )

```

For the main analyses we have four fixed effects of interest. (1) The effect of need fulfillment on interaction quality: _b_ = `r round(fixef(lmer_quality_contemporary)['keymotive_fulfillemt_1_cwc'], 3)`, (2) the effect of need fulfillment on outgroup attitudes: _b_ = `r round(fixef(lmer_attitude_contemporary)['keymotive_fulfillemt_1_cwc'], 3)`, (3) the effect of need fulfillment on outgroup attitudes while controlling for interaction quality: _b_ = `r round(fixef(lmer_attitude_need_quality_contemporary)['keymotive_fulfillemt_1_cwc'], 3)`, as well as (4) the effect of interaction quality on outgroup attitudes while controlling for need fulfillment: _b_ = `r round(fixef(lmer_attitude_need_quality_contemporary)['quality_overall_1_cwc'], 3)`.

# Power Simulation Parameters

```{r}
#| label: simulation setup
nsim <- 1000
progress <- FALSE
n_ppt <- 250
n_measure <- 180 # 3 months with 2 measurements per day

# extrapolate data:
sim_n_quality <- extend(lmer_quality_contemporary, along="PID", n = n_ppt)
sim_n_attitude <- extend(lmer_attitude_contemporary, along="PID", n = n_ppt)
sim_n_attitude_full <- extend(lmer_attitude_need_quality_contemporary, along="PID", n = n_ppt)

# extrapolate data:
sim_t_quality <- extend(lmer_quality_contemporary, within="PID", n = n_measure)
sim_t_attitude <- extend(lmer_attitude_contemporary, within="PID", n = n_measure)
sim_t_attitude_full <- extend(lmer_attitude_need_quality_contemporary, within="PID", n = n_measure)
```

@@

## Power of `Quality ~ NeedFulfill` for different numbers of participants

```{r}
#| label: participant simulation Quality NeedFulfill
#| warning: FALSE
#| error: FALSE

# Run and plot simulations
pc_n_quality_need = powerCurve(
  fit = sim_n_quality,
  test = simr::fixed("keymotive_fulfillemt_1_cwc", "t"),
  along = "PID",
  progress = progress,
  breaks = seq(1, 12, 1),
  nsim = nsim
)
plot(pc_n_quality_need) +
  title("Quality given Need Fulfillment [participant numbers]")
```

## Power of `Attitude ~ NeedFulfill` for different numbers of participants

```{r}
#| label: participant simulation Attitude NeedFulfill
#| warning: FALSE
#| error: FALSE

# Run and plot simulations
pc_n_attitude_need = powerCurve(
  fit = sim_n_attitude,
  test = simr::fixed("keymotive_fulfillemt_1_cwc", "t"),
  along = "PID",
  progress = progress,
  breaks = seq(1, 45, 1),
  nsim = nsim
)
plot(pc_n_attitude_need) +
  title("Attitudes given Need Fulfillment [participant numbers]")
```

## Power of `Attitude ~ NeedFulfill + Quality` for different numbers of participants

we expect this to be super small. Keep in mind that the _b_ = `r round(fixef(lmer_attitude_need_quality_contemporary)["keymotive_fulfillemt_1_cwc"],3)`. So even with a large number of participants, we would not expect these effects to be meaningful.

```{r}
#| label: participant simulation Attitude NeedFulfill given Quality
#| warning: FALSE
#| error: FALSE

# Run and plot simulations
pc_n_attitude_need = powerCurve(
  fit = sim_n_attitude_full,
  test = simr::fixed("keymotive_fulfillemt_1_cwc", "t"),
  along = "PID",
  progress = progress,
  breaks = seq(10, 250, 10),
  nsim = nsim
)
plot(pc_n_attitude_need) +
  title("Attitudes given Need Fulfillment (controling for Quality) [participant numbers]")
```

## Power of `Attitude ~ Quality` given NeedFulfill for different numbers of participants

```{r}
#| label: participant simulation Attitude Qaulity given NeedFulfill
#| warning: FALSE
#| error: FALSE

# Run and plot simulations
pc_n_attitude_full_quality = powerCurve(
  fit = sim_t_attitude_full,
  test = simr::fixed("quality_overall_1_cwc", "t"),
  along = "PID",
  progress = progress,
  breaks = c(seq(1, 25, 1), seq(30, 100, 10)),
  nsim = nsim
)
plot(pc_n_attitude_full_quality) +
  title("Attitudes given Quality (controlling for Need Fulfillment) [participant numbers]")
```

@@

## Power of `Quality ~ NeedFulfill` for different numbers of measurements

```{r}
#| label: measurement simulation Quality NeedFulfill
#| warning: FALSE
#| error: FALSE

# Run and plot simulations
pc_t_quality_need = powerCurve(
  fit = sim_t_quality,
  test = simr::fixed("keymotive_fulfillemt_1_cwc", "t"),
  within = "PID",
  progress = progress,
  breaks = seq(1, 20, 1),
  nsim = nsim
)
plot(pc_t_quality_need) +
  title("Quality given Need Fulfillment [measurement numbers]")
```

## Power of `Attitude ~ NeedFulfill` for different numbers of measurements

```{r}
#| label: measurement simulation Attitude NeedFulfill
#| warning: FALSE
#| error: FALSE

# Run and plot simulations
pc_t_attitude_need = powerCurve(
  fit = sim_t_attitude,
  test = simr::fixed("keymotive_fulfillemt_1_cwc", "t"),
  within = "PID",
  progress = progress,
  breaks = seq(1, 25, 1),
  nsim = nsim
)
plot(pc_t_attitude_need) +
  title("Attitudes given Need Fulfillment [measurement numbers]")
```

## Power of `Attitude ~ NeedFulfill` given Quality for different numbers of measurements

we expect this to be super small. Keep in mind that the _b_ = `r round(fixef(lmer_attitude_need_quality_contemporary)["keymotive_fulfillemt_1_cwc"],3)`. So even with a large number of measurements, we would not expect these effects to be meaningful.

```{r}
#| label: measurement simulation Attitude NeedFulfill given Quality
#| warning: FALSE
#| error: FALSE

# Run and plot simulations
pc_t_attitude_full_need = powerCurve(
  fit = sim_t_attitude_full,
  test = simr::fixed("keymotive_fulfillemt_1_cwc", "t"),
  within = "PID",
  progress = progress,
  breaks = seq(10, 180, 10),
  nsim = nsim
)
plot(pc_t_attitude_full_need) +
  title("Attitudes given Need Fulfillment (controlling for Quality) [measurement numbers]")
```

## Power of `Attitude ~ Quality` given NeedFulfill for different numbers of measurements

```{r}
#| label: measurement simulation Attitude Qaulity given NeedFulfill
#| warning: FALSE
#| error: FALSE

# Run and plot simulations
pc_t_attitude_full_quality = powerCurve(
  fit = sim_t_attitude_full,
  test = simr::fixed("quality_overall_1_cwc", "t"),
  within = "PID",
  progress = progress,
  breaks = seq(1, 15, 1),
  nsim = nsim
)
plot(pc_t_attitude_full_quality) +
  title("Attitudes given Quality (controlling for Need Fulfillment) [measurement numbers]")
```

@@

# save simulations, because this takes for freakin' ever.      
save(list = ls(pattern = "plt."), 
     file = "data/S1_Workers/processed/PowerSimPlots.RData")

>>>>

# Install and Load Packages
# !IMPORTANT!
# BEFORE FIRST RENDER:
# To install all relevant packages please run "renv::restore()" (or renv::init() and then initiate from lockfile) in the console before the first use to ensure that all packages are using the correct version.
# to store the packages in a contained library within the project folder: renv::settings$use.cache(FALSE) and add 'RENV_CONFIG_SANDBOX_ENABLED = FALSE' to an '.Renviron' file
lib <- c(
  "rmarkdown",
  "knitr",
  "remedy",
  "bookdown",
  "brms",
  "psych",
  "ggplot2",
  "ggthemes",
  # "haven",
  "RColorBrewer",
  # "plotly",
  "gridExtra",
  "ggpattern",
  "ggridges",
  "binom",
  "iterators",
  "pbkrtest",
  "plotrix",
  "RLRsim",
  "stats",
  "methods",
  "utils",
  "graphics",
  "grDevices",
  "car",
  "testthat",
  "simr",
  "data.table",
  "lme4",
  "lmerTest",
  # "nlme",
  # "jtools",
  # "gtsummary",
  "sessioninfo",
  # "tibble",
  "pander",
  # "devtools",
  # "mada",
  #"tidyr",
  #"tidyverse",
  "plyr",
  "dplyr",
  # "Hmisc",
  "kableExtra",
  # "papaja",
  "stringr"#,
  # "stringi",
  # "reshape2",
  # "lubridate",
  # "purrr",
  # "metafor"
)
invisible(lapply(lib, library, character.only = TRUE))
rm(lib)

@@

Generally, power considerations of mixed effects model such as repeated measures ANCOVAs and multi-level regressions are difficult to estimate because of the complex covariance structures. We ran simulation-studies based on our initial results following the first study. We found that sufficient sample sizes heavily depended on the type and size of the effect. When we simulated models with the marginalization sub-scale (which was the best distributed with mostly small to medium effect sizes), we found that for between-subjects effects, a sample size of 80 participants would have been sufficient (power above .8) with two measurements (see Figure 9 for an illustration). However, for the same between-subject effects we did not reach a power above .8 with any increase of repeated follow-up measurements (based on model extrapolations with a fixed _n_ = 20; tested up to 12 months). The within time effect, on the other hand, was estimated well above .8 by the seventh measurement (i.e., after six months, see Figure 9). The cross-level interaction effects (within-between) were largely a mixture of their individual effects and changes in measurement numbers and sample size affected the models accordingly. We, therefore, urge researchers to rely on pilot data and to run extensive simulation studies to maximize effort and feasibility (i.e., sufficient power).

@@

load("data/S1_Workers/processed/df.btw.RData")
```

@@

# prepare data for regression (i.e, put into long format and change variable types)
marginalization = melt(df.btw[!is.na(df.btw$assimilation.post),c("marginalization.pre","marginalization.post",
                                                                 "CtContactNL_c","AvKeyNeedInt_c",
                                                                 "ExternalReference")],
                       id=c("ExternalReference","CtContactNL_c","AvKeyNeedInt_c"))
marginalization$time = ifelse(grepl(".pre", marginalization$variable), "pre", "post")
marginalization$variable = gsub(".pre$|.post$", "", marginalization$variable)
marginalization$time = factor(marginalization$time)
marginalization$time = factor(marginalization$time,levels(marginalization$time)[c(2,1)])
marginalization$id = rep(seq(1:20), 2)
```

@@

# run regresion 
lm.marg <- lmer(value~
                  AvKeyNeedInt_c*time+
                  CtContactNL_c*time+
                  (1|id), 
                data=marginalization)
summary(lm.marg)
fixef(lm.marg)
```

# Post-Hoc Power

Check post-hoc power for comparison.

```{r}
#| label: post hoc simulation

# Post-hoc power analysis for single effects (example: key need)
post_hoc <- powerSim(lm.marg, fixed("AvKeyNeed"), nsim=1000)
post_hoc
```

@@

## Power of contact effect for different numbers of participants

```{r}
#| label: participant simulation contact

# extrapolate data:
sim.N <- extend(lm.marg, along="id", n=500)

# Run and plot simulations
plt.pwr.contact = powerCurve(fit = sim.N, 
                             test = simr::fixed("CtContactNL_c", "t"), 
                             along="id", 
                             progress = FALSE,
                             breaks = seq(10,100,10), 
                             nsim=1000)
plot(plt.pwr.contact) +
  title("Outgroup Contact Count [participant numbers]")
```

## Power of time effect for different numbers of participants

```{r}
#| label: participant simulation time

plt.pwr.time = powerCurve(sim.N, simr::fixed("time"), 
                          along="id", 
                          progress = FALSE, 
                          breaks = seq(10,100,10), 
                          nsim=1000)
plot(plt.pwr.time) +
  title("Time [participant numbers]")
```

## Power of need fulfillment effect for different numbers of participants

```{r}
#| label: participant simulation avg need fulfillment

plt.pwr.Need = powerCurve(sim.N, simr::fixed("AvKeyNeedInt_c", "t"), 
                          along="id", 
                          progress = FALSE, 
                          breaks = seq(10,100,10), 
                          nsim=1000)
plot(plt.pwr.Need) +
  title("Avergage Need Fulfillment [participant numbers]")
```

## Power of time by contact effect for different numbers of participants

```{r}
#| label: participant simulation contact by time

plt.pwr.timeCont = powerCurve(sim.N, simr::fixed("time:CtContactNL_c"), 
                             along="id", 
                             progress = FALSE, 
                             breaks = seq(10,500,50), 
                             nsim=1000)
plot(plt.pwr.timeCont) +
  title("Time X Outgroup Contact Count [participant numbers]")
```

## Power of time by need fulfillment effect for different numbers of participants

```{r}
#| label: participant simulation need fulfillment by time

plt.pwr.timeNeed = powerCurve(sim.N, simr::fixed("AvKeyNeedInt_c:time"), 
                              along="id", 
                              progress = FALSE, 
                              breaks = seq(10,50,5), 
                              nsim=1000)
plot(plt.pwr.timeNeed) +
  title("Time X Avergage Need Fulfillment [participant numbers]")
```

@@

## Power of time effect for different numbers of measurements

```{r}
#| label: measurement simulation time

# extrapolate data:
sim.within <- extend(lm.marg, within="id", n=12)

# Run and plot simulations
plt.pwr.t.time = powerCurve(sim.within, simr::fixed("time"), 
                          within="id", 
                          progress = FALSE, 
                          breaks = seq(2,12,1), 
                          nsim=1000)
plot(plt.pwr.t.time) +
  title("Time [measurement numbers]")
```

## Power of contact effect for different numbers of measurements

```{r}
#| label: measurement simulation contact count

plt.pwr.t.contact = powerCurve(sim.within, simr::fixed("CtContactNL_c", "t"), 
                               within="id", 
                               progress = FALSE, 
                               breaks = seq(2,12,1), 
                               nsim=1000)
plot(plt.pwr.t.contact) +
  title("Outgroup Contact Count [measurement numbers]")
```

## Power of need fulfillment effect for different numbers of measurements

```{r}
#| label: measurement simulation contact need fulfillment

plt.pwr.t.Need = powerCurve(sim.within, simr::fixed("AvKeyNeedInt_c", "t"), 
                            within="id", 
                            progress = FALSE, 
                            breaks = seq(2,12,1), 
                            nsim=1000)
plot(plt.pwr.t.Need) +
  title("Average Need Fulfillment [measurement numbers]")
```

## Power of time by contact effect for different numbers of measurements

```{r}
#| label: measurement simulation contact time by contact count

plt.pwr.t.timeCont = powerCurve(sim.within, simr::fixed("time:CtContactNL_c"), 
                          within="id", 
                          progress = FALSE, 
                          breaks = seq(2,12,1), 
                          nsim=1000)
plot(plt.pwr.t.timeCont) +
  title("Time X Outgroup Contact Count [measurement numbers]")
```

## Power of time by need fulfillment effect for different numbers of measurements

```{r}
#| label: measurement simulation need fulfillment by time

plt.pwr.t.timeNeed = powerCurve(sim.within, simr::fixed("AvKeyNeedInt_c:time"), 
                               within="id", 
                               progress = FALSE, 
                               breaks = seq(2,12,1), 
                               nsim=1000)
plot(plt.pwr.t.timeNeed) +
  title("Time X Average Need Fulfillment [measurement numbers]")
```

@@

# save simulations, because this takes for freakin ever.      
save(plt.pwr.contact,plt.pwr.Need,plt.pwr.t.contact,plt.pwr.t.Need,
     plt.pwr.t.time,plt.pwr.t.timeCont,plt.pwr.t.timeNeed,plt.pwr.time,
     plt.pwr.timeCont,plt.pwr.timeNeed, 
     file = "data/S1_Workers/processed/PowerSimPlots.RData")
